---
title: "Regex"
author: "Yukthi"
date: "29/01/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r imports}
library(dplyr)
MData = read.csv("measurements.csv")
```


# Modification of Species Column

In this file I re-read the measurements.csv file into a datafam called MData. I then produced a new column by using the dplyr.transmute function to return a dataframe column using the sub function to change the "Species" values to the new truncated form. After doing this I then assigned this returned column to a new column in the MData dataframe called Sp.

Finally using the select function I made a new dataframe called newData which took all columns in the MData dataframe but replaced the Species column with the Sp column.

```{r manipulation}
MData= mutate(MData, Sp = sub('(\\w)\\w+',
                         '\\1\\.',
                         as.vector(t(MData['Species'])))
                        )
newData = select(MData,Sp,Limb.Width:unitsL)
print(newData)
```
